---
title: "EAF calculation process"
author: "Electricity Authority"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(here)
source(here::here('libraries.R'))

Sys.setenv('EMI_SAS' = "sv=2019-12-12&si=exp2022-03-31&sr=c&sig=m2kSNI61k%2BmyHl1O84EdFlaXbgRMve%2BNByuDIzoMMXg%3D")
offers_from_trade_date='2016-01-01'
offers_to_trade_date='2016-12-31'
from_trade_date='2016-01-01'
to_trade_date='2016-12-31'
scenario_type=2
run_name='default'
output_path='data/output'
```

# Overview

This vignette shows an example of the EAF calculation process. This takes you through generating the offer adjustments, configuring a vSPD experiment for the counterfactual, and calculating the final EAF from the results. 

# Setup

Follow the instructions from the README to install all necessary dependencies. The SAS you configure for access to EMI datasets should have the format shown.

```{r}
Sys.getenv('EMI_SAS')
```

# Generating offer overrides for scenario 2

First source `main.R` to load the necessary functions. To generate the scenario 2 adjustment  for 2016 then run.

```{r, eval = FALSE}
source('main.R')

generate_offer_adjustments(offers_from_trade_date = '2016-01-01',
                           offers_to_trade_date = '2016-12-31',
                           run_name = "default",
                           scenario_type = '2',
                           dl_offers = TRUE,
                           dl_assumptions = TRUE)
```

Note that we only need to download input offers and assumptions once. You can set `dl_offers` and `dl_assumptions` to `FALSE` on subsequent runs to speed things up.

This can then be edited to produce an offer adjustment override for different periods.

This then produces an output file ``r file.path(output_path, 'GDX', glue("{offers_from_trade_date}-{offers_to_trade_date}-scenario{scenario_type}-{run_name}-ets-offer-overides.gdx"))``, which is used with vSPD. You will need to move this to the directory `{path to your vSPD installation}/Override/`.

# Running counterfactual simulations

You will need a GAMS license to run this simulation. First download the final pricing case files for the same period as in stage 1.

```{r, eval = FALSE}

download_emi_final_pricing_case_files(from_trade_date = '2019-01-01',
                                      to_trade_date = '2019-12-31',
                                      download_dir = "{path to your vSPD installation}/Input")

```

Then generate the file list for the counterfactual experiment.

```{r, eval = FALSE}

generate_file_manifest_dir(in_dir = '{path to your vSPD installation}/Input'
                           out_file = '{path to your vSPD installation}/Programs/vSPDfileList.inc')

```

Finally edit the settings file `{path to your vSPD installation}/vSPDSettings.inc`, the `vSPDinputOvrdData` path should be set to the name of your override file in the overide folder (e.g. ``r glue("{offers_from_trade_date}-{offers_to_trade_date}-scenario{scenario_type}-{run_name}-ets-offer-overides.gdx")``). You may also wish to set the `runName` option. 

Once this is configured run `gams runvSPD` to run the experiment. You will need the file `{runName}_NodeResults_TP.csv` from the results to calculate EAF factors. You may wish to move this from the results to the default input directory `data/input/prices/Counterfactual/`.

# Generating offer adjustments for scenario 3 

Using the price file from the previous step, we can then calculate the scenario 3 offer adjustment. 

```{r, eval = FALSE}
generate_offer_adjustments(offers_from_trade_date = '2016-01-01',
                           offers_to_trade_date = '2016-12-31',
                           run_name = "default",
                           scenario_type = '3',
                           df_cf_prices_file = glue('data/input/prices/Counterfactual/{runName}_NodeResults_TP.csv'),
                           dl_offers = TRUE,
                           dl_assumptions = TRUE)
```

This then gives us a new override file for the same time period. You can repeat the counterfactual simulation step with this override file to produce the scenario 3 counterfactual prices.

# Generating EAF factors

Finally we calculate the EAF for the period of interest. Note that this currently only supports annual results.

```{r, eval = FALSE}
calculate_annual_eaf(from_trade_date,
                     to_trade_date,
                     cf_prices_file,
                     scenario_type = c('2', '3'),
                     run_name = "",
                     dl_fp = TRUE,
                     dl_assumptions = TRUE)
```

As in the offer adjustment, we only require `dl_fp` and `dl_assumptions` to be `TRUE` the first time we run a particular calculation.
This will then output a file (by default ``r glue("data/output/eaf/{from_trade_date}-{to_trade_date}-scenario-{scenario_type}-{run_name}-eaf.csv")``) containing the EAF for the period of interest.

